A comparison of design-based and model-based approaches for spatial data.

### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
model-based background revision.

On the other hand, model-based inference imposes additional assumptions on the data with a potential to provide more precise estimators if the additional assumptions hold. Instead of estimating fixed but unknown parameters, the goal of model-based inference in the spatial context is often \emph{prediction} of an unknown quantity. This is a fundamental philosophical difference between sampling-based and model-based approaches. Instead of \emph{estimating} a fixed unknown mean, we are \emph{predicting} the value of the mean, a random variable. We know that if we sampled all sites, we would have an exact prediction for the mean of our one realized spatial surface, without any uncertainty. But, the true mean of the spatial process that generated our realized data is still not known, and, in the prediction context, we typically do not care much about what value the mean of the underlying process takes.

>>>>

The model-based approach assumes the data are a random realization of a process. Randomness is often incorporated through distributional assumptions on the data-generating process. Instead of estimating fixed but unknown parameters (as in the design-based approach), the goal of model-based inference in the spatial context is often \emph{prediction} of an unknown quantity. For example, suppose the realized mean of all population units is the quantity of interest. Instead of \emph{estimating} a fixed unknown mean, we are \emph{predicting} the value of the mean, a random variable. We know that if we sampled all population units, we would have an exact prediction for the mean of our one realized process, without any uncertainty. But the true mean of the spatial process that generated our realized data is still not known. When predicting the realized mean, we typically are not interested in the underlying process' true mean.

Assuming the data is a realization of a specific data-generating process yields predictors that are linked to distributional assumptions. These distributional assumptions are used to derive prediction intervals. The distributional assumptions allow the prediction intervals can be more precise. @Cressie2015statistics and @Schabenberger2017statistical provide reviews of model-based approaches for spatial data.
